---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(arrow)
library(modelsummary)
library(bayesplot)
library(broom)
library(gridExtra)
library(rstanarm)
library(knitr)
```


# Introduction

You can and should cross-reference sections and sub-sections. 

The remainder of this paper is structured as follows. @sec-data....



# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-sample
#| fig-cap: Bills of penguins
#| echo: false

analysis_data <- read_parquet(here::here("outputs/data/analysis_data.parquet"))
kable(head(analysis_data))
```


```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

distribution_price <-
  ggplot(analysis_data, aes(x = Price)) +
  geom_histogram() +
  theme_minimal()

distribution_EPS <-
  ggplot(analysis_data, aes(x = EPS)) +
  geom_histogram() +
  theme_minimal()

distribution_Dividends <-
  ggplot(analysis_data, aes(x = Dividends)) +
  geom_histogram() +
  theme_minimal()

distribution_Net_Income <-
  ggplot(analysis_data, aes(x = Net_Income)) +
  geom_histogram() +
  theme_minimal()

grid.arrange(distribution_price, distribution_EPS, distribution_Net_Income, distribution_Dividends, nrow = 2)
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| include:false

summary(analysis_data$Price)
summary(analysis_data$EPS)
summary(analysis_data$Net_Income)
summary(analysis_data$Dividends)
```


```{r}
#| label: tbl
#| fig-cap: Relationship between me and ur mom
#| echo: false
#| warning: false
#| message: false

base_net_income_plot <-
  analysis_data |> 
  ggplot(aes(x = Net_Income, y = Price)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(x = "Net Income",
       y = "Stock Price")

base_dvidends_plot <-
  analysis_data |> 
  ggplot(aes(x = Dividends, y = Price)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(x = "Dividends",
       y = "Stock Price")

base_eps_plot <-
  analysis_data |> 
  ggplot(aes(x = EPS, y = Price)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(x = "EPS",
       y = "Stock Price")

best_fit_net_income <-
  base_net_income_plot +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = "y ~ x"
  )

best_fit_dividends <-
  base_dvidends_plot +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = "y ~ x"
  )

best_fit_eps <-
  base_eps_plot +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = "y ~ x"
  )

grid.arrange(best_fit_net_income, best_fit_dividends, best_fit_eps, ncol = 2, nrow = 2)

```



Talk way more about it. 



# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

Simple Model
\begin{align} 
y &= \beta_0 + \beta_1X + \epsilon \\
Y &\sim \mbox{Normal}(\beta, \sigma^2)
\end{align}
Where Y is the stock price and X is earnings per share.

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

Multivariable Model
\begin{align} 
y &= \beta_0 + \beta_1X_1 + \beta_2X_2 - \beta_3X_3 + \epsilon \\
Y &\sim \mbox{Normal}(\beta, \sigma^2)
\end{align}
Where X1 is EPS, X2 is net income, and X3 is dividends.

### Model justification
We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

Our results are summarized in @tbl-modelresults.



```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

first_model <-
  lm(
    Price ~ EPS,
    data = analysis_data
  )

modelsummary(list("Net_Income" = first_model))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-poopy but hole
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

analysis_data <-
  augment(
    first_model,
    data = analysis_data
  )

distribution_residuals <-
  ggplot(analysis_data, aes(x = .resid)) +
  geom_histogram(binwidth = 1) +
  theme_classic() +
  labs(y = "Number of occurrences", x = "Residuals")

residuals_eps <-
  ggplot(analysis_data, aes(x = EPS, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
  theme_classic() +
  labs(y = "Residuals", x = "Earnings Per Share")

residuals_price <-
  ggplot(analysis_data, aes(x = Price, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
  theme_classic() +
  labs(y = "Residuals", x = "Stock Price")

estimated_vs_actual <-
  ggplot(analysis_data, aes(x = Price, y = .fitted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_classic() +
  labs(y = "Estimated Stock Price", x = "Actual Stock Price")

grid.arrange(distribution_residuals, residuals_eps, residuals_price, estimated_vs_actual, nrow = 2)

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

  ggplot(analysis_data,aes(y=Price,x=EPS,color=Net_Income)) +
  geom_point() +
  stat_smooth(method="lm",se=FALSE)+
  theme_classic() 

```


```{r}
#| echo: false
#| eval: true
#| label: tbl-uuhhhh
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

multi_model <-
  lm(
    Price ~ EPS + Net_Income,
    data = analysis_data
  )

modelsummary(
  list(
    "EPS Only" = first_model,
    "With Net Income" = multi_model
  ),
  fmt = 2
)
```

```{r}
#| include: false
#| eval: true
#| warning: false
model_rstanarm_1 <-
  stan_glm(
    formula = Price ~ EPS + Net_Income,
    data = analysis_data,
    family = gaussian(),
    prior = normal(location = 0, scale = 2.5),
    prior_intercept = normal(location = 0, scale = 2.5),
    prior_aux = exponential(rate = 1),
    seed = 345
  )

saveRDS(
  model_rstanarm_1,
  file = "model_rstanarm_1.rds"
)
```


```{r}
#| include: false
#| eval: true
#| warning: false

model_rstanarm_1 <-
  readRDS(file = "model_rstanarm.rds")
prior_summary(model_rstanarm_1)
```

```{r}
#| include: false
#| eval: true
#| warning: false
model_rstanarm_2 <-
  stan_glm(
    formula = Price ~ EPS + Net_Income,
    data = analysis_data,
    family = gaussian(),
    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
    prior_intercept = normal(0, 2.5, autoscale = TRUE),
    prior_aux = exponential(rate = 1, autoscale = TRUE),
    seed = 1234
  )

saveRDS(
  model_rstanarm_2,
  file = "model_rstanarm_2.rds"
)
```


```{r}
#| echo: false
#| eval: true
#| label: tbl-poop
#| tbl-cap: "priors"
#| warning: false

modelsummary(
  list(
    "Non-scaled priors" = model_rstanarm_1,
    "Auto-scaling priors" = model_rstanarm_2
  ),
  fmt = 2
)
```


```{r}
prior_summary(model_rstanarm_2)
```









# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(model_rstanarm_2) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(model_rstanarm_2) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```


## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...


```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2
plot(model_rstanarm_2, "trace")

plot(model_rstanarm_2, "rhat")
```

\newpage


# References


